Our Galactic Midplane Is Probably Wrong, and Not by a Little

### 2019-01-29 paper/ms.tex
More small updates to text

\begin{figure*}
\plotone{fig/schmactions_one_orbit.pdf}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick disk
orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
and $z_{\text{max}}=425\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
observe the orbit over $1\,\Gyr$ and perform the same orbit integration
procedure at each timestep, and report the values of the actions. The
computation of $L_z$ is pristine to errors in $z$, with only numerical
artifacts remaining. Only small errors are induced in $J_r$, with the middle
$90\%$ of values over the $\Gyr$ being $\sim6\%$ of the true $J_r$. As
expected, large errors are induced in $J_z$, with the middle $90\%$ of values
being $\sim85\%$ of the true $J_z$.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

We now repeat the procedure but continuously changing the error in the $z$ and
$x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth
minus 5\uth percentile divided by the true action value for three different
orbits. The \Gus{orange} orbit is exactly the same as in
Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider two more planar
orbits with initial vertical velocities of $50\,\kms$ (\Gus{red}) and
$10\,\kms$ (\Gus{blue}). For these two new orbits, the true actions are $(J_r,
L_z, J_z) = (32.6, -1520, 22.9)\,\actunit$ and $(40.4, -1520, 0.7)\,\actunit$
for the \Gus{red} and \Gus{blue} orbits, respectively. The value of
$z_{\text{max}}$ for these orbits are $854\,\pc$ and $122\,\pc$, respectively
(with the orange orbit having $z_{\text{max}}$ of $425\,\pc$.) \Gus{Easy way
to explain change in $J_r$?}

@@

The {\em upper right} panel shows that the fractional error in $J_z$ is more
exaggerated for more planar orbits. For the nearly planar (blue) orbit,
deviations that are $\sim30\,\pc$ already give $100\%$ deviations in the
actions. The required offset for $100\%$ deviation is $\sim120$ and
$\sim225\,\pc$ for the orange and red orbits, respectively. \Gus{Make these
numbers exact.}

>>>>

\begin{figure*}
\plotone{fig/schmactions_one_orbit.pdf}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick disk
orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
observe the orbit over $1\,\Gyr$ and perform the same orbit integration
procedure at each timestep, and report the values of the actions. The
computation of $L_z$ is pristine to errors in $z$, with only numerical
artifacts remaining. Only small errors are induced in $J_r$, with the middle
$90\%$ of values over the $\Gyr$ being $\sim8\%$ of the true $J_r$. As
expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
with the middle $90\%$ of values being $\sim43\%$ of the true $J_z$. The $x$
offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and $\sim3\%$,
respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

We now repeat the procedure but continuously changing the error in the $z$ and
$x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth
minus 5\uth percentile divided by the true action value for three different
orbits. The thick (brown) orbit is exactly the same as in
Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider the thin and halo
orbits (pink and blue, respectively) from Table~\ref{tab:orbits}.

@@

The {\em upper right} panel shows that the fractional error in $J_z$ is more
exaggerated for more planar orbits. For the thin orbit, deviations that are
$\sim30\,\pc$ already give $100\%$ deviations in the actions. The required
offset for $100\%$ deviation is $\sim225\,\pc$ for the thick orbit. The halo
orbit is relatively resistant to errors in the midplane, with only $\sim25\%$
error in $J_z$ out to a midplane offset of $500\,\pc$.
